---
title: "Coursera"
author: "Gunther E"
date: "8/25/2021"
output: html_document
---
Q1 loading and processing the data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(dplyr)
library(ggplot2)

# cleaning the date class
activity$date <- ymd(activity$date)

#changing activity df name to data
data <- activity

```

## R Markdown
Q2: What is mean total number of steps taken per day?

In this section I will summarize the total number of step taken per day

```{r Summarize}

summary(data)

#without NA data
summarize <- data   %>%
  group_by(date) %>%
   summarize(
  steps_per_day = sum(steps))

```
Getting the mean and median for the steps per day
```{r, mean and median}
mean <- as.integer(mean(summarize$steps_per_day))
median<- as.integer(median(summarize$steps_per_day))
```

## Including Plots

Here I make a histogram of the steps per day

```{r pressure, echo=FALSE}

ggplot(data=summarize, aes(steps_per_day)) + 
  geom_histogram(fill = "darkblue", color = "white") + ggtitle("Steps per day") + theme_dark()

```

The user of this step counter walked 10,000 steps per day the most. 



##############################################

Q3: What is the average daily activity pattern?
```{r, make plot with median and mean}

steps_interveal <- data %>% group_by(interval) %>% summarise(mean_steps = mean(steps, 
    na.rm = TRUE))

 ggplot(data = steps_interveal, mapping = aes(x = interval, y = mean_steps)) + 
    geom_line() + scale_x_continuous("daily intervel", breaks = seq(min(steps_interveal$interval), 
    max(steps_interveal$interval), 100)) + scale_y_continuous("Average Number of Steps") + 
    ggtitle("Average Number of Steps Taken by Interval")
```
Starting around 0800 we see the largest peak. The user of the step counter is rarely walking before 5:30 or after 10:00. So we could possibly assume the user is not moving much from 1000 to 0500. 


###################################################

Q4: Imputing missing values

```{r}
#looking at the na data, number
sum(is.na(data))

#what does the data look like without na
omit <- na.omit(activity)

#adding data to NA values
add <- activity
add$steps[is.na(add$steps)] <- mean(add$steps,na.rm=TRUE)
add$steps <- as.numeric(add$steps)
add$interval <- as.numeric(add$interval)
colSums(is.na(add))

#with NA data replaced
summarize_with <- add   %>%
  group_by(date) %>%
   summarize(
  steps_per_day = sum(steps))

#gettung the new mean and median
mean_add <- as.integer(mean(summarize_with$steps_per_day))
median_add<- as.integer(median(summarize_with$steps_per_day))
```
```{r}

#making a plot with the new data set 
ggplot(data=summarize_with, aes(steps_per_day)) + 
  geom_histogram(fill = "darkblue", color = "white") + ggtitle("Steps per day - with NA") + theme_dark()

```

It seems that there is not much differnce between the two data sets besides that there are now more steps in the 9000 range


######################################################

Q5:
Are there differences in activity patterns between weekdays and weekends?

```{r}

add$day <- ifelse(weekdays(add$date) %in% c("Saturday","Sunday"), "weekday", "weekend")

# Preparing data for ggplot
week_data <- add %>% group_by(interval,day) %>% summarise(mean=mean(steps))

# Plot Average steps across weekday/weekend vs 5-min interval Time Series
ggplot(week_data, aes(x=interval, y=mean, color=day)) + 
  facet_wrap(day~.) +
  geom_line(color = "darkblue") + 
  labs(title="Average number of step compared by weekday vs weekend", y="Average Number of Steps", x="Time series") + theme_grey()

```
There seem to be differences in between weekday and weekends. During the week the user walks more than during the weekend. The user of the walk monitor also seems to start walking early in the day on average on the weekends. Perhaps this means that the user wakes up for an early morning walk on the weekends but not during the weekdays. However, in both cases the user walks the most from around 8 o'clock to 9. 
